---
title: "Reproducible Research Assignment 1"
author: "mac"
date: "4 novembre 2014"
output: html_document
---
**What is mean total number of steps taken per day?**

Make an histogram of the total number of steps taken each day.
```{r}
setwd("/Users/mac/Documents/BigData/datasciencecoursera/ReproRes")
activ<-read.csv("activity.csv",sep=",")
# Prepare somewhere to store the values of interest
with(activ,tapply(steps,date, sum,na.rm=T)) ->sums
# Plotting a barplot with the collected values
barplot(sums,xlab="Date",ylab="Steps",main="Steps per day")
```

Calculate and Report the mean and the median total number of steps taken per day 
```{r}
print (mean(sums,na.rm = T))
print (median(sums,na.rm = T))
```
**What is the average daily activity pattern?**
```{r}
#Calculating means for each interval
with(activ,tapply(steps,interval, mean,na.rm=T)) ->interval_means

plot(unique(activ$interval),interval_means,type="l",ylab="number of steps",xlab="intervals")
#Retrieving the maximum of the interval means
index_interest<-which.max(interval_means)
#Retrieving the corresponding index
print(activ$interval[[index_interest]])
```
**Imputing missing values**
Print the number of rows with NAs
```{r}
indexes <- which(is.na(activ))
print (nrow(activ[indexes,]))
```
Devise a strategy for filling in all of the missing values in the dataset.

```{r}
#Creating a new column with the corresponding interval means
activ<-cbind(activ,rep(interval_means,61))
#adding the right name on the 4th column
colnames(activ) <-c("steps","date","interval","interval_means")
# inject interval means values on the right positions
activ[indexes,]$steps <- activ[indexes,]$interval_means
# We calculate a new sum
with(activ,tapply(steps,date, sum,na.rm=T)) ->sums2
# Plotting a barplot with the collected values
barplot(sums2,xlab="Date",ylab="Steps",main="Steps per day")
print (mean(sums2,na.rm=T))
print (median(sums2,na.rm=T))
```
After having imputing the mean number of steps in missing values, the mean and the median have shifted and are now equal.

**Are there differences in activity patterns between weekdays and weekends?**
```{r}
#Creating the weekends variable which determine wether the date is in a weekend
weekends <- weekdays(as.Date(activ$date)) %in% c("Samedi","Dimanche")

library(lattice)
#Adding the variable to the data frame
activ<-cbind(activ,weekends)
print(head(activ))
#Aggregating infos in a new data frame (newdf)
aggregate(steps~interval+weekends,activ,mean)  -> newdf

#Plotting the data
xyplot(steps~interval|weekends,data=newdf,type="l",main="Average number of steps taken ",ylab="Number os steps",xlab="Intervals",strip=strip.custom(factor.levels=c("WeekDays","WeekEnd")))

```


During week ends the number of steps reaches a higher maximum than in week days. On the other hand, the average number of steps seems to be higher.


